---
title: "Exploratory Analysis"
author: "Jacob Ye, Matt Taniguchi (Add your names here)"
date: "11/19/2021"
---

```{r include=FALSE}
library("knitr")
library("ggplot2")
library("ggforce")

source("aggregate_table.R")
source("chart1.R")
source("chart2.R")
```

# Introduction
In this analysis, we are looking at a life expectancy dataset obtained from Kaggle.
This data is originally sourced from the WHO, and contains information regarding
life expectancy and other factors which affect it (education, alcohol consumption, 
rates of diseases, etc.).

# Summary
The set contains data from **193 different countries** in the **years 2000 to 2015**.
There are **22 features** in this set with **2938 observations**. There are a
total of **161 developing** countries and **32 developed** countries. The ratio 
between the developing to developed countries and the ratio of developing to 
developed observations matches closely. Both developing and developed countries 
share a **maximum life expectancy of 89**, while the developing countries' minimum 
is almost half **(36.3)** of the developed countries minimum **(69.9)**. This 
trend of developing countries having the lower end of life expectancies was expected.
Through further analysis, we will be able to understand which factors in these 
developing countries have a negative effect on life expectancy.

# Aggregate Table

```{r echo=FALSE}
kable(aggregate_table)
```

The most meaningful way to group each country was by whether it was developing or
developed, which is why we only have two rows in this table. The averages are simply
the means of the corresponding values (expectancy, population, gdp). Despite the
small number of rows in this table, we can still obtain a lot of information about 
the average differences between developing and developed countries. 
Developing countries have a **noticeably lower life expectancy**
than developed countries, with much **higher populations** and **significantly less GDP**
than developed countries. It is astonishing to see how larger developing countries'
populations are, almost **twice that of developed countries**. Even more astonishing but
somewhat unsurprising is how developed countries' average GDP almost **quintuples 
that of developing countries**.

# Chart 1
## Life Expectancy over Time

```{r echo=FALSE}
# Chart 1
expectancy_time_plot
```

We included this chart to show how life expectancy has changed over the years.
We think that it's important to see the general trend and see if it
aligns of what we think how life expectancy should change, and We think that
it does corroborate with our beliefs that as time passes on our health as
technology gets better. Though it is important to mention that despite knowing
that it improves over time, it doesn't show what is causing it explicity to
increase.

# Chart 2
## aaaaaa
chart 2 here

# Chart 3
## aaaaaa
chart 3 here

***
